---
title: "Author R Package Functions That Extend The Ready4 Framework"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Author R Package Functions That Extend The Ready4 Framework}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4fun)
```

`ready4fun` provides classes, methods and functions to support highly standardised approaches to authoring R packages that extend the [ready4 framework](https://www.ready4-dev.com/) for open and modular mental health systems models.

`ready4fun` builds on existing development workflow tools (e.g. `devtools`) to ensure that ready4 framework R packages are:

- *citable* (with a Zenodo generated DOI and an algorithm generated CITATION file);
- *community-friendly* (applying deprecation conventions supported by `lifecycle`); 
- *documented* (applying a function self-documenting algorithm that extends `sinew`, deploying a GitHub pages hosted and `pkgdown` generated website and authoring PDF manuals stored in a GitHub Release via `piggyback`);
- *internally consistent* implementing automated checks to ensure consistency in naming conventions, etc; 
- *licensed* (via a `usethis` generated GPL-3 license); 
- *quality assured* (using continuous integration via GitHub actions and R-CMD-Check); and
- *versioned* (applying `usethis` version increments).

The main class exported as part of ready4fun is `ready4fun_manifest` in which the metadata for an R package is declared. The two main functions from `ready4fun` that are most relevant to end users are `make_pkg_desc_ls` and `make_manifest` as using both of these functions together provides the optimal means of creating a valid instance of the `ready4fun_manifest` class. The `ready4fun_manifest` class and `make_pkg_desc_ls` and `make_manifest` functions are designed to be used as part of [the workflow supported by the ready4pack package](https://ready4-dev.github.io/ready4pack/articles/V_01.html).

## Future documentation

It is important to note that the workflows supported by `ready4fun` functions are designed to be implemented after an empty R package has been created with undocumented functions, authored using the `ready4fun` house-style, that are saved in sub-directories of your package's `data-raw` directory. Guidance for how to apply the house style has yet to be published, but more detailed documentation on using `ready4fun` will be created in 2022.
