---
title: "A Standardised Workflow For Authoring R Packages"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{A Standardised Workflow For R Package Creation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4fun)
```

`ready4fun` provides classes, methods and functions to support highly standardised approaches to authoring R packages that extend the [ready4 framework](https://www.ready4-dev.com/) for open and modular mental health systems models.

ready4fun builds on `devtools` and other development workflow tools to ensure that the ready4 framework R packages authored with it are:

- *citable* (with a Zenodo generated DOI and an algorithm generated CITATION file);
- *community-friendly* (applying deprecation conventions supported by `lifecycle`); 
- *documented* (applying a function self-documenting algorithm that extends `sinew`, deploying a GitHub pages hosted and `pkgdown` generated website and authoring PDF manuals stored in a GitHub Release via `piggyback`);
- *internally consistent* implementing automated checks to ensure consistency in naming conventions, etc; 
- *licensed* (via a `usethis` generated GPL-3 license); 
- *quality assured* (using continuous integration via GitHub actions and R-CMD-Check); and
- *versioned* (applying `usethis` version increments).

The main class exported as part of ready4fun is `ready4fun_manifest` in which the metadata for an R package is declared. In most cases, the methods associated with `ready4fun_manifest` do not need to be called directly by a user. Instead they are now incorporated in the methods of the [ready4pack](https://ready4-dev.github.io/ready4pack/) package. The two main functions from `ready4fun` that are most relevant to end users are `make_pkg_desc_ls` and `make_manifest` as using both of these functions together provides the optimal means of creating a valid instance of the `ready4fun_manifest` class.

The workflows supported by `ready4fun` functions are designed to be implemented after an empty R package has been created with draft functions, authored using the `ready4fun` house-style, saved in sub-directories of the package `data-raw` directory. Detailed instructions on applying the house style have yet to be released, but examples of how `ready4fun` functions integrate into package authoring workflows using `ready4pack` are available. 

These examples can be found in the DATASET file within the data-raw folder of any R package that extends the ready4 framework (see https://www.ready4-dev.com/toolkits/). For example, to see how `ready4fun_manifest` is used in the workflow to author the [ready4show](https://ready4-dev.github.io/ready4show/) package, view [this file in the source code](https://github.com/ready4-dev/ready4show/blob/main/data-raw/DATASET.R).

A more detailed guide to using ready4fun will be created in 2022.
