---
title: "Workflow Tools To Make 'Under The Hood' Code More Transparent"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Workflow Tools To Make 'Under The Hood' Code More Transparent}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4fun)
library(ready4use)
library(ready4show)
```

## Motivation

The [ready4 framework](https://www.ready4-dev.com/) for open and modular mental health systems models is [implemented using an object-oriented programming (OOP) approach](https://ready4-dev.github.io/ready4class/articles/V_01.html). One motivation for using OOP is the concept of "abstraction" - making things as simple as possible for end-users of `ready4` "methods" (algorithms) by exposing the minimal amount of code required to implement each method.

However, some users of the `ready4` framework will want to "look under the hood" and examine the code that implements methods in much more detail. Reasons to do so include to:

 - gain detailed insight into how methods are implemented;
 - test individual sub-components ("functions") of methods as part of code verification and model validation checks;
 - re-use functions when authoring new methods.

Therefore when authoring code libraries that extend the `ready4` framework, it is important to ensure that "under the hood" code can be readily understood. Two ways for achieving this goal is to ensure that all functions (even those not intended for use by modeller end-users) are adequately documented and adopt a consistent house style (e.g. naming conventions). ready4fun provides workflow tools (classes, methods, functions and datasets) to achieve these goals.

## ready4fun function authoring taxonomies, abbreviations and workflow

The `ready4fun` package uses a dataset of taxonomies and abbreviations to ensure standardised function code style and documentation. A copy of this dataset (`x`) can be downloaded from a repository associated with the [ready4 package](https://ready4-dev.github.io/ready4/) using tools from the [ready4use package](https://ready4-dev.github.io/ready4use/) package.

```{r}
x <- ready4use::Ready4useRepos(gh_repo_1L_chr = "ready4-dev/ready4",
                               gh_tag_1L_chr = "Documentation_0.0") %>%
  ingest(metadata_1L_lgl = F)
```

### Function names begin with a meaningful verb

Consistent with a naming convention popular in the R development community, all `ready4` framework functions begin with a verb. Furthermore, the choice of verb is meaningful - it communicates something about the type of task a function implements. For example, all functions beginning with the word "fit" will fit a model of a specified type to a dataset. The definitions of all meaningful verbs currently used by ready4 functions (excluding methods) are stored in element `fn_types_lup` of `x`, the key features of which are reproduced below.

```{r fnstb, tab.cap='Meaningful verbs', tab.id = 'fnstb', results="asis"}
x$fn_types_lup %>% 
  dplyr::filter(!is_generic_lgl &!stringr::str_detect(fn_type_nm_chr,pattern = " ")) %>% dplyr::select(1:2) %>%
  dplyr::rename(Verb = fn_type_nm_chr,
                Description = fn_type_desc_chr) %>%
  ready4show::print_table(caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          output_type_1L_chr = "HTML",
                          mkdn_tbl_ref_1L_chr = knitr::opts_current$get("tab.id"))
```

### Function inputs and outputs have meaningful suffices

The type of input (arguments) required and output (return) produced by a function can be efficiently communicated by using meaningful suffices. For example all objects ending in "_chr" are character vectors and all objects ending in "_int" are integer vectors. The meaningful suffices currently used by to describe objects in the ready4 framework are stored in element `seed_obj_lup_tb` of `x`, the key features of which are reproduced below.

```{r sfcstb, tab.cap='Meaningful suffices', tab.id = 'sfcstb', results="asis"}
x$seed_obj_lup_tb %>% 
dplyr::select(1:2) %>%
  dplyr::rename(Suffix = short_name_chr,
                Description = long_name_chr) %>%
  ready4show::print_table(caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          output_type_1L_chr = "HTML",
                          mkdn_tbl_ref_1L_chr = "fnstb")
```

### Consistent use of abbreviations

Further information about the purpose of a function and the nature of its inputs and outputs can be encoded by using naming conventions that make consistent use of abbreviations. A master table of the abbreviations used throughout the `ready4` framework is maintained in the `abbreviations_lup` element of `x`. The list of abbreviations is now quite extensive and continues to grow as the `ready4` suite of software expands. The initial few entries of `abbreviations_lup` are reproduced below.

```{r abbs, tab.cap='Abbreviations', tab.id = 'abbs', results="asis"}
x$abbreviations_lup %>% 
dplyr::select(1:2) %>%
  dplyr::rename(Abbreviation = short_name_chr,
                Description = long_name_chr) %>%
  head() %>%
  ready4show::print_table(caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          output_type_1L_chr = "HTML",
                          mkdn_tbl_ref_1L_chr = "fnstb")
```

### Workflow

The main class exported as part of ready4fun is `ready4fun_manifest` which is used to specify metadata (including details of the repository in which the `fn_types_lup`, `seed_obj_lup_tb` and `abbreviations_lup` objects are stored) for an R package that is being authored. A `ready4fun_manifest` object is created with the aid of the `make_pkg_desc_ls` and `make_manifest` functions and is designed to be used as part of [the ready4pack R package authoring workflow package](https://ready4-dev.github.io/ready4pack/articles/V_01.html). 

That vignette includes links to two examples of where the `ready4pack` workflow has been used to author R package. To illustrate how `readyfun` tools used as part of that workflow are used to document functions, we are just going to focus on [the program used to create the ready4show package](https://github.com/ready4-dev/ready4show/blob/main/data-raw/DATASET.R).

That program makes use of `ready4fun` tools that read all [undocumented package functions](https://github.com/ready4-dev/ready4show/tree/main/data-raw/fns), performs automated checks to ensure that these functions appropriately use the taxonomies and abbreviations mentioned previously (prompting authors to make specific amendments if they do not) and then rewrites these functions to the [package R directory](https://github.com/ready4-dev/ready4show/tree/main/R), appending tags (with the aid of the `sinew` package) that will generate meaningful documentation. 

For example, one of the functions to be documented is the [knit_from_tmpl](https://github.com/ready4-dev/ready4show/blob/main/data-raw/fns/knit.R), which is transformed to a [version with tags](https://github.com/ready4-dev/ready4show/blob/main/R/fn_knit.R). The tags added to all functions are then used to generate the package documentation, including the package manual. Two versions of the `ready4show` package manual are generated - [a slimmed down version for end-users](https://github.com/ready4-dev/ready4show/releases/download/Documentation_0.0/ready4show_User.pdf) and [a more detailed inventory of contents intended for developers](https://github.com/ready4-dev/ready4show/releases/download/Documentation_0.0/ready4show_Developer.pdf).


## Future documentation

Detailed guidance for how to apply `ready4fun` workflow tools has yet to be prepared but will be released in 2022.
