---
title: "A Standardised Workflow For Authoring R Packages"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{A Standardised Workflow For R Package Creation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(ready4fun)
```

ready4fun provides classes, methods and functions to support highly standardised approaches to authoring R packages that extend the ready4 framework for open and modular mental health systems models.

The main class exported as part of ready4fun is `ready4fun_manifest` in which the metadata for an R package is declared. In most cases, the methods associated with `ready4fun_manifest` do not need to be called directly by a user. Instead they are now incorporated in the methods of the [ready4pack](https://ready4-dev.github.io/ready4pack/) package. 

Examples of how `ready4fun_manifest` is intended to be integrated into package authoring workflows using `ready4pack` can be found in the DATASET file within the data-raw folder of any R package that extends the ready4 framework (see https://www.ready4-dev.com/toolkits/). For example, to see how `ready4fun_manifest` is used in the workflow to author the [ready4use](https://ready4-dev.github.io/ready4use/) package, view [this file in the source code](https://github.com/ready4-dev/ready4use/blob/master/data-raw/DATASET.R).

A more detailed guide to using ready4fun will be created in 2022.
